%%
%% This is file `thesis-ex.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% uiucthesis2009.dtx  (with options: `example')
%% 
\def\fileversion{v2.25a} \def\filedate{2009/10/10}
%% Package and Class "uiucthesis2009" for use with LaTeX2e.
\documentclass[edeposit,10pt,fullpage]{uiucthesis2009}

\usepackage{amsmath, amsthm, amssymb}
\usepackage[numbers]{natbib}

\usepackage{color}
\usepackage{verbatim}

\usepackage{subfigure}
\usepackage{url}
\usepackage{times}
\usepackage{mathptmx}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[pdftex]{graphicx}
\usepackage{booktabs}

\usepackage{subfigure}
\usepackage{caption}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning,fit,scopes}
\usetikzlibrary{shapes.multipart}

\colorlet{d}{black!25}
\colorlet{l}{black!0}
\newcommand{\btnode}[4]{ \nodepart{#1} #2\\ $[$#3,#4) }
\newcommand{\btplainnode}[2]{ \nodepart{#1} #2  \\  }
\tikzset{nd/.style={rectangle split, rectangle split parts=3, draw=black!75, thick,
            rectangle split horizontal, align=center, rectangle split empty part width=0.5cm,
            rectangle split part fill={#1}}}

\usepackage[pdfborder={0 0 0}]{hyperref}

\interfootnotelinepenalty=10000  % Split footnote are annoying

\newcommand{\ntaskip}{\vspace{-0.15in}}
\SetAlgoSkip{ntaskip}

\newenvironment{captiontext}{%
   \begin{center}%
     \begin{minipage}{0.99\linewidth}%
       \renewcommand{\baselinestretch}{0.9}%
         \sffamily\footnotesize}%
   {\renewcommand{\baselinestretch}{1.0}%
      \end{minipage}%
        \end{center}}

\newenvironment{smitemize}%
  {\begin{list}{$\bullet$}%
     {\setlength{\parsep}{0pt}%
      \setlength{\topsep}{0pt}%
      \setlength{\itemsep}{2pt}}}%
  {\end{list}}

\clubpenalty=10000  % Don't allow orphans
\widowpenalty=10000 % Don't allow widows

% \raggedbottom

\renewcommand{\textfraction}{0}
\renewcommand{\topfraction}{1}
\renewcommand{\bottomfraction}{1}
\setcounter{totalnumber}{10}
\setcounter{topnumber}{10}
\setcounter{bottomnumber}{10}
\setcounter{dbltopnumber}{10}
\renewcommand{\floatpagefraction}{1}
\renewcommand{\dblfloatpagefraction}{1}

\begin{document}

\title{Storage System Design for Non-Volatile Byte-Addressable memory using Consistent and Durable Data Structures}
\author{Shivaram Venkataraman}
\department{Computer Science}
\schools{M.S, University of Illinois at Urbana-Champaign, 2011}
\msthesis
\advisor{Roy H. Campbell}
\degreeyear{2011}
\committee{Professor Roy H. Campbell}
\maketitle

\frontmatter

%% Create an abstract that can also be used for the ProQuest abstract.
%% Note that ProQuest truncates their abstracts at 350 words.
\begin{abstract}
The predicted shift to low cost, non-volatile, byte-addressable memory (e.g.,
Phase Change Memory and Memristor), the growth of ``big data'', and the
subsequent emergence of frameworks such as memcached and NoSQL systems require
us to rethink the design of storage systems.  Today, the predominant method
available to application programmers for obtaining persistence of data is
through the file interface which provides access to a filesystem, which in turn
uses the block layer to provide access to hard disk drives or solid state
disks.  With the advent of non-volatile byte-addressable memory (NVBM),
buffered I/O systems designed primarily for hard disks are no longer the most
efficient design for storage.  We show that the overheads imposed by system
calls, the file interface, and the block device layer can make NVBM access up
to 266 times slower compared to directly performing loads and stores.

With performance characteristics closer to DRAM, NVBM demands software designs
that resemble common heap data structures rather than current systems that
serialize data via the file interface. To derive the maximum performance, we
propose the use of single-level data stores where no distinction is made
between a volatile and a persistent copy of data. For these systems, we present
Consistent and Durable Data Structures (CDDSs) that, on current hardware,
allows programmers to safely exploit the low-latency and non-volatile aspects
of new memory technologies.  CDDSs use versioning to allow atomic updates
without requiring logging.  The same versioning scheme also enables rollback
for failure recovery.  When compared to a memory-backed Berkeley DB B-Tree, our
prototype-based results show that a CDDS B-Tree can increase \textit{put} and
\textit{get} throughput by~74\% and 138\%. Tembo, a CDDS B-Tree enabled
distributed Key-Value system, has up to~250\%--286\% higher throughput than
Cassandra, a two-level data store.

%When compared to Cassandra, a
%two-level data store, Tembo, a CDDS B-Tree enabled distributed Key-Value
%system, increases throughput by up to~250\%--286\%.

%We contend that software designed for NVBM will bear a closer resemblance
%to common heap data structures running on top of DRAM than current systems
% that serialize data via the file interface.

\end{abstract}

%% Create a dedication in italics with no heading, centered vertically
%% on the page.
\begin{dedication}
To my\\
Mother and Father
\end{dedication}

%% Create an Acknowledgements page, many departments require you to
%% include funding support in this.
\chapter*{Acknowledgments}

I would like to thank my advisor Roy H. Campbell for his
invaluable guidance over the course of my Masters degree and for giving me
the opportunity to be a part of the Systems Research Group. This has helped
me learn a lot about academic research.  I am deeply indebted to Niraj
Tolia for encouraging me to work on the design of storage systems for
non-volatile memory and much of the work described in this thesis was
born out of our discussions. I would also like to thank Matthew Caesar for
his encouragement and Parthasarathy Ranganathan, Nathan L. Binkert
for their valuable inputs towards this work.

I would like to thank Ellick Chan, Abhishek Verma and Reza Farivar for working
with me on multiple research projects over the last two years. I am also thankful to 
Alejandro Gutierrez, Mirko Montanari, Cristina Abad, Kevin Larson and Will Dietz
for the many technical discussions, and all the members of Systems Research Group
for the vibrant research environment.

My research has been funded by a fellowship from the Siebel Scholars Foundation
and internships at HP Labs, Palo Alto. I would also like to thank Cinda Heeren
for the invaluable skills I learnt and the wonderful experience I had as a
Teaching Assistant. Lynette Llubben, the staff at the Computer Science
department and the staff at HP Labs have helped me with many administrative
tasks and I thank them for their help.  My time at Urbana-Champaign has been
made wonderful by support from many friends and family. Finally I would like to
thank my parents for their constant support and for encouraging me to
pursue my dreams.

%% The thesis format requires the Table of Contents to come
%% before any other major sections, all of these sections after
%% the Table of Contents must be listed therein (i.e., use \chapter,
%% not \chapter*).  Common sections to have between the Table of
%% Contents and the main text are:
%%
%% List of Tables
%% List of Figures
%% List Symbols and/or Abbreviations
%% etc.

\tableofcontents
\listoftables
\listoffigures
\listofalgorithms

\addcontentsline{toc}{chapter}{List of Algorithms}
%% Create a List of Abbreviations. The left column
%% is 1 inch wide and left-justified
%\chapter{List of Abbreviations}
%
%\begin{symbollist*}
%\item[CA] Caffeine Addict.
%\item[CD] Coffee Drinker.
%\end{symbollist*}

%% Create a List of Symbols. The left column
%% is 0.7 inch wide and centered
%\chapter{List of Symbols}

%\begin{symbollist}[0.7in]
%\item[$\tau$] Time taken to drink one cup of coffee.
%\item[$\mu$g] Micrograms (of caffeine, generally).
%\end{symbollist}

\mainmatter
\include{chapters/intro}
\include{chapters/background}
\include{chapters/fs}
\include{chapters/btree-design}
\include{chapters/kv}
\include{chapters/eval}
\include{chapters/profiling}
\include{chapters/conclusion}

\appendix*

%\include{Appendix.tex}

\backmatter

\bibliographystyle{abbrvnat}
%\bibliographystyle{latex8}
\bibliography{ref}

\end{document}
\endinput
%% End of file `thesis-ex.tex'.
